---
output: github_document
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
execute:
  echo: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE
)

library(sf)
library(tidyverse)
library(tmap)
```

# wasteskipsblantyre

<!-- badges: start -->
[![R-CMD-check](https://github.com/openwashdata/wasteskipsblantyre/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/openwashdata/wasteskipsblantyre/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

# Overview

The wasteskipsblantyre data package includes locations of the publicly accessible waste skips in Blantyre, Malawi. The data was collected in 2021.

## Installation

You can install the development version of wasteskipsblantyre from [GitHub](https://github.com/) with:

```{r echo=TRUE, eval=FALSE}
# install.packages("devtools")
devtools::install_github("openwashdata/wasteskipsblantyre")
```

Alternatively, you can download the individual datasets as a CSV or XLSX
file from the table below.

```{r, echo=FALSE}

extdata_path <- "wasteskipsblantyre/tree/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |> 
  distinct(file_name) |> 
  mutate(file_name = str_remove(file_name, ".rda")) |> 
  rename(dataset = file_name) |> 
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |> 
  knitr::kable()

```

# Project

## TITLE_OF_PROJECT

(One sentence) Evaluating the potential of ........

### Description

(One to two short paragraphs) This Project focuses on determining
.......

### Research Question

(One to two questions)

### Data

The data set inlcludes the locations of the publicly accessible waste skips in Blantyre, Malawi. In the context of the **PROJECT_NAME** project described above the data was collected in **MONTH** of 2021.

The package provides access to one data set.

```{r, echo = TRUE}
library(wasteskipsblantyre)
```

The `wasteskipsblantyre` data set has `r ncol(wasteskipsblantyre)` variables and `r nrow(wasteskipsblantyre)` observations. For an overview of the variable names, see the following table.  

```{r, eval=FALSE}
wasteskipsblantyre
```

```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |> 
  dplyr::select(variable_name:description) |> 
  knitr::kable()
```


```{r, echo=FALSE}
#| label: fig-waste-skips-simple
#| fig-cap: A small visualisation of this data set

sf_wsb <- st_as_sf(wasteskipsblantyre, coords = c('long', 'lat'), crs = 4326)
tmap_mode("plot")
sf_wsb |>
  tm_shape() +
  tm_dots() +
  tm_graticules()
```

## Example

This is a basic example which shows you how to use the data:

```{r example, eval=FALSE, echo=TRUE}
library(wasteskipsblantyre)
library(sf)
library(tmap)

# read data set into a simple feature (spatial vector data)
sf_wsb <- st_as_sf(wasteskipsblantyre, coords = c('long', 'lat'), crs = 4326)
# set mapping mode to interactive ("view")
tmap_mode("view")
# create an interactive map 
qtm(sf_wsb)
```
