---
output: github_document
editor_options:
  chunk_output_type: console
  markdown:
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)

library(sf)
library(tidyverse)
library(wasteskipsblantyre)
```

# wasteskipsblantyre

<!-- badges: start -->

[![DOI](https://zenodo.org/badge/483225869.svg)](https://zenodo.org/badge/latestdoi/483225869)
[![R-CMD-check](https://github.com/openwashdata/wasteskipsblantyre/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/openwashdata/wasteskipsblantyre/actions/workflows/R-CMD-check.yaml)
[![License: CC BY 4.0](https://img.shields.io/badge/License-CC_BY_4.0-lightgrey.svg)](https://creativecommons.org/licenses/by/4.0/)

<!-- badges: end -->

The goal of wasteskipsblantyre is to provide data for research and
planning of solid waste management in Blantyre, Malawi. The package
contains two datasets: `skips` holds the locations of the publicly
accessible waste skips in the city, with the number of skips, access
and the fill-up times reported at each location, and `poi` holds the
points of interest for collecting the waste from the skips. The data
were updated in 2023 and replace the 2021 dataset `wasteskipsblantyre`
of earlier versions.

```{r map-skips, echo=FALSE}
wards <-
  st_read(here::here("vignettes", "articles",
                     "traditional-authorities", "Malawi_TA_2018.shp"),
          quiet = TRUE) |>
  filter(DIST_NAME == "Blantyre City")

sf_skips <- skips |>
  filter(!is.na(x), !is.na(y)) |>
  st_as_sf(coords = c("x", "y"), crs = 4326)

sf_poi <- st_as_sf(poi, coords = c("long", "lat"), crs = 4326)

ggplot() +
  geom_sf(data = wards, fill = "grey95", colour = "grey60") +
  geom_sf(data = sf_skips,
          aes(size = fct_na_value_to_level(factor(number_skips),
                                           level = "not recorded")),
          colour = "#5b195b", alpha = 0.7) +
  scale_size_manual(values = c("0" = 1.5, "1" = 4, "2" = 6,
                               "not recorded" = 2.5)) +
  geom_sf(data = sf_poi, shape = 17, size = 3, colour = "#e08214") +
  labs(title = "Public waste skips in Blantyre, Malawi",
       subtitle = "Skip locations (circles) and points of interest (triangles), 2023",
       size = "Number of skips") +
  theme_void()
```

## Installation

You can install the development version of wasteskipsblantyre from
[GitHub](https://github.com/) with:

```{r eval=FALSE, echo=TRUE}
# install.packages("devtools")
devtools::install_github("openwashdata/wasteskipsblantyre")
```

Alternatively, you can download the individual datasets as a CSV or XLSX
file from the table below.

```{r, echo=FALSE}

extdata_path <- "https://github.com/openwashdata/wasteskipsblantyre/raw/main/inst/extdata/"

read_csv("data-raw/dictionary.csv") |>
  distinct(file_name) |>
  mutate(file_name = str_remove(file_name, ".rda")) |>
  rename(dataset = file_name) |>
  mutate(
    CSV = paste0("[Download CSV](", extdata_path, dataset, ".csv)"),
    XLSX = paste0("[Download XLSX](", extdata_path, dataset, ".xlsx)")
  ) |>
  knitr::kable()

```

## Project goal

Data on the location of public waste skips in Blantyre was not
available. Without such data, it becomes difficult to develop a solid
waste management plan and logistics. The goal of this project was to
identify the locations of public waste skips in Blantyre, Malawi, and
to record how they are used.

## Data

The package provides access to the locations of the publicly accessible
waste skips in Blantyre, Malawi, and to the points of interest for their
collection.

```{r, echo = TRUE}
library(wasteskipsblantyre)
```

### skips

The `skips` dataset has `r ncol(skips)` variables and `r nrow(skips)`
observations. `r sum(!is.na(skips$name))` rows are named skip locations,
`r sum(!is.na(skips$x))` of them with coordinates. The last
`r sum(is.na(skips$name))` rows have no name, no location and no other
values; they carry only the constant capacity of 7000 litres per skip
(see [issue
#11](https://github.com/openwashdata/wasteskipsblantyre/issues/11)). For
an overview of the variable names, see the following table.

```{r, eval=FALSE, echo=TRUE}
skips
```

```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "skips.rda") |>
  dplyr::select(variable_name:description) |>
  knitr::kable()
```

### poi

The `poi` dataset has `r ncol(poi)` variables and `r nrow(poi)`
observations: the depot, the Mzedi dump site, the compost site and two
gas stations. For an overview of the variable names, see the following
table.

```{r, eval=FALSE, echo=TRUE}
poi
```

```{r, echo=FALSE}
readr::read_csv("data-raw/dictionary.csv") |>
  dplyr::filter(file_name == "poi.rda") |>
  dplyr::select(variable_name:description) |>
  knitr::kable()
```

## Example

The code below is an example which shows how you could use the data to
prepare an interactive map in R. Find more examples in the [prepared
examples
article](https://openwashdata.github.io/wasteskipsblantyre/articles/examples.html).

```{r example, eval=FALSE, echo=TRUE}
library(wasteskipsblantyre)
library(dplyr)
library(sf)
library(tmap)

# read the skip locations with coordinates into a simple feature
sf_skips <- skips |>
  filter(!is.na(x), !is.na(y)) |>
  st_as_sf(coords = c("x", "y"), crs = 4326)
# set mapping mode to interactive ("view")
tmap_mode("view")
# create an interactive map
qtm(sf_skips)
```

## License

Data are available as
[CC-BY](https://github.com/openwashdata/wasteskipsblantyre/blob/main/LICENSE.md).

## Citation

Please cite using:

```{r}
citation("wasteskipsblantyre")
```
