---
title: "Examples"
author: Sebastian Camilo Loos
editor_options: 
  chunk_output_type: console
output: 
  bookdown::html_document2:
    fig_caption: yes
    toc: true
    toc_depth: 2
    number_sections: FALSE
pkgdown:
  as_is: true
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r unknown_chunk, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  message = FALSE,
  warning = FALSE,
  fig.retina = 2,
  fig.align = 'center'
)
```

::: callout-note
**The following article is an example that shows how to work with the data package and produce insights using R.**
:::

## Setting things up

### Load Packages

```{r setup, echo = TRUE, output = FALSE, message = FALSE, warning = FALSE}
# install.packages("devtools")
# devtools::install_github("openwashdata/wasteskipsblantyre")
library(wasteskipsblantyre)
library(tidyverse)
library(sf)
library(here)
```

```{r scrape-datasets, eval=FALSE, include=FALSE}

# # get traditional authorities boundaries
# # download and unzip maps from https://data.humdata.org
# utils::download.file(url = "<https://data.humdata.org/dataset/9f3896ae-d77b-414a-85ef-8ede33389ed0/resource/ac6dfe69-3e06-4b16-907f-18d5674ecbc7/download/mwi_ta_adm3_nso_2018_20190327.zip>",
#                      destfile = here::here("vignettes", "articles",
#                                            "traditional-authorities.zip"),
#                      mode = "wb")
# unzip(here::here("vignettes", "articles", "traditional-authorities.zip"), exdir = here::here("vignettes", "articles", "traditional-authorities"))
# 
# # remove zip file
# file.remove(here::here("vignettes", "articles", "traditional-authorities.zip"))

# # download and unzip cenusus data from https://data.humdata.org
# utils::download.file(url = "https://data.humdata.org/dataset/9f3896ae-d77b-414a-85ef-8ede33389ed0/resource/8e3dcdeb-a2e6-4f84-aa3f-5430761b88b6/download/2018_census_total_number_of-hh-per-ea-csv.csv",
#                      destfile = here::here("vignettes", "articles",
#                                            "traditional-authorities", 
#                                            "census_data.csv"),
#                      mode = "wb")
```

### Read in Data Sets

First, additional files including data sets with specific census data about Blantyre has been added to the package from <https://data.humdata.org>. For this fictional article we will use spatial data and therefore, the data has to be read in as [simple features](https://en.wikipedia.org/wiki/Simple_Features) using the [`sf-package`](https://r-spatial.github.io/sf/) as you can see in the following code chunk.

```{r read-data, echo=TRUE, results='hide'}
# read in ward boundaries data
census_boundaries <- 
  st_read(here::here("vignettes", "articles",
                     "traditional-authorities", "Malawi_TA_2018.shp")) |> 
  st_as_sf()

# read in traditional authority population data
census_data <- read_csv(here::here("vignettes", "articles", 
                                   "traditional-authorities", 
                                   "census_data.csv"))

# read the skip locations with coordinates into a simple feature
# (spatial vector data) and set the coordinate reference system equal to
# the one from census_boundaries
sf_skips <- skips |> 
  filter(!is.na(x), !is.na(y)) |> 
  st_as_sf(coords = c("x", "y"), crs = 4326) |> 
  st_transform(st_crs(census_boundaries))

# wards of Blantyre City
wards <- census_boundaries |> 
  filter(DIST_NAME == "Blantyre City")
```

## Background

Malawi is an eastern sub-Saharan state. For this article the area of the city of Blantyre has been explored as depicted in the following map.

```{r boundaries-blantyre, fig.cap="Boundaries of the wards in Blantyre."}
ggplot(wards) +
  geom_sf(fill = "grey95", colour = "grey40") +
  theme_void()
```

The city is divided in `r nrow(wards)` wards. This article will use these areas to explore the data contained in the `wasteskipsblantyre` package. The `skips` dataset stores the locations of publicly accessible waste skips in Blantyre (map below), updated in 2023. `r sum(!is.na(skips$x))` of its `r nrow(skips)` rows have coordinates; the rows without coordinates are left out of this article.

```{r locations, fig.cap="Locations of the publicly accessible waste skips in Blantyre, Malawi."}
ggplot() +
  geom_sf(data = wards, fill = "grey95", colour = "grey40") +
  geom_sf(data = sf_skips, colour = "#5b195b", size = 2) +
  theme_void()
```

In addition, data from the 2018 Malawi Census population data set shared by National Statistical Office is used for some demographic analysis.

## Hypothesis

In the context of this fictional article we define a hypothesis:

*The number of waste skips within traditional authorities (wards) correlates with the population of each authority.*

## Analysis

```{r join-data, include=FALSE}
# count the skips (number_skips) in each ward of Blantyre City
count_ta <- st_join(wards, sf_skips) |> 
  group_by(TA_NAME) |> 
  summarise(n = sum(number_skips, na.rm = TRUE)) |> 
  mutate(mean_number = mean(n))

pop_ta <- census_data |> 
  filter(DISTRICT == "Blantyre City") |> 
  summarise(ta_pop = sum(TOTAL_POP), .by = TA_NAME)

density_data <- count_ta |> 
  left_join(pop_ta, by = "TA_NAME") |> 
  mutate(area_km2 = as.numeric(st_area(count_ta) / 1000000),
         density_popkm2 = ta_pop / area_km2)

no_skip_wards <- density_data |> 
  st_drop_geometry() |> 
  filter(n == 0) |> 
  arrange(desc(density_popkm2)) |> 
  pull(TA_NAME)
```

First, we have a separate look into the waste skips data and the data from the 2018 census for each ward within Blantyre. After that we compare the two data sets and look for possible correlations.

### Waste Skips Data

The `skips` dataset records `r sum(skips$number_skips, na.rm = TRUE)` public waste skips at `r sum(!is.na(skips$name))` named locations; some locations have no skip anymore or no count of skips. In Figure \@ref(fig:count-ta) it is visible that the number of waste skips in a ward varies between `r min(count_ta$n)` and `r max(count_ta$n)`. On average, there are `r round(mean(count_ta$n), 1)` waste skips in a ward.

```{r count-ta, fig.cap="Number of waste skips in each ward."}
count_ta |> 
  ggplot(aes(y = forcats::fct_rev(TA_NAME), x = n)) +
  geom_col(fill = "#5b195b") +
  geom_vline(xintercept = mean(count_ta$n)) +
  labs(x = "Number of publicly accessible waste skips",
       y = "Ward in Blantyre") +
  theme_bw()
```

The city center is located to the west of the city. The map below colours each ward by the number of waste skips within its boundaries.

```{r map-count, fig.cap="Wards coloured according to the number of waste skips that they have within their boundaries."}
count_ta |> 
  mutate(n = factor(n)) |> 
  ggplot() +
  geom_sf(aes(fill = n), colour = "grey40", alpha = 0.8) +
  geom_sf(data = sf_skips, size = 1) +
  scale_fill_brewer(palette = "RdPu") +
  labs(fill = "Number of publicly\naccessible waste skips") +
  theme_void()
```

### Population Data

```{r pop-blantyre}
pop_blantyre <- sum(pop_ta$ta_pop)
```

In 2018, Blantyre City's population was `r format(pop_blantyre, big.mark = ",")`. The population for each ward can be seen in Figure \@ref(fig:barplot-pop).

```{r barplot-pop, fig.cap="Population per ward."}
pop_ta |> 
  ggplot(aes(x = ta_pop, y = forcats::fct_rev(TA_NAME))) +
  geom_col(fill = "#5b195b") +
  labs(x = "Ward Population",
       y = "Ward in Blantyre") +
  theme_bw()
```

Since the areas of the wards differ, we will have a look into the population density of the wards in Figure \@ref(fig:barplot-density).

```{r barplot-density, fig.cap="Population density per ward."}
density_data |> 
  ggplot(aes(x = density_popkm2, y = forcats::fct_rev(TA_NAME))) +
  geom_col(fill = "#5b195b") +
  labs(x = "Population Density (per km^2)",
       y = "Ward in Blantyre") +
  theme_bw()
```

### Comparison

In order to cross check the data and validate the hypothesis the data sets are linked together and visualized in the following figures.

Figure \@ref(fig:count-pop) shows the population of each ward, coloured by the number of waste skips within the ward. The correlation between the number of waste skips and the ward population is `r round(cor(density_data$n, density_data$ta_pop), 2)`.

```{r count-pop, fig.cap="Ward population, coloured by the number of waste skips within the ward."}
density_data |> 
  ggplot(aes(x = ta_pop, y = forcats::fct_rev(TA_NAME))) +
  geom_col(aes(fill = factor(n))) +
  scale_fill_brewer(palette = "RdPu") +
  labs(x = "Ward Population",
       y = "Ward in Blantyre",
       fill = "Number of publicly\naccessible waste skips") +
  theme_bw()
```

Figure \@ref(fig:count-density) does the same for the population density. The correlation between the number of waste skips and the population density is `r round(cor(density_data$n, density_data$density_popkm2), 2)`. The wards without a waste skip are, from the most to the least densely populated: `r knitr::combine_words(no_skip_wards)`.

```{r count-density, fig.cap="Ward population density, coloured by the number of waste skips within the ward."}
density_data |> 
  ggplot() +
  geom_col(aes(x = density_popkm2, y = forcats::fct_rev(TA_NAME), fill = factor(n))) +
  scale_fill_brewer(palette = "RdPu") +
  labs(x = "Population Density (per km^2)",
       y = "Ward in Blantyre",
       fill = "Number of publicly\naccessible waste skips") +
  theme_bw()
```

## Conclusion

The correlation coefficients above show how closely the number of waste skips in a ward follows its population and population density. Check the [source code](https://github.com/openwashdata/wasteskipsblantyre/blob/main/vignettes/articles/examples.Rmd) to reproduce the analysis with the `skips` data.

::: callout-note
**We hope you enjoyed this fictional article!**
:::
